---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all()
pkgs <- get_all_packages()
```

# The Cognitive Effects of Computational Thinking: A Systematic Review and Meta-Analytic Study

<!-- badges: start -->
[<img alt="alt_text" src="https://img.shields.io/badge/OSF-DOI%2010.17605%2FOSF.IO%2FUVBCD-blue" />](https://osf.io/uvbcd/)
<!-- badges: end -->

This repository contains all the scripts, functions and data to fully reproduce the statistical analysis of the *The cognitive effects of computational thinking: A systematic review and meta-analytic study* paper by Montuori C., Gambarota F., Altoè G. and Arfè B.

For a detailed description of the analysis method see the **supplementary materials** ([HTML](docs/supplementary/supplementary.html), [PDF](docs/supplementary/supplementary.pdf))

## Setup

In order to reproduce the analysis is necessary to clone or download this repository and open the `metaCoding.Rproj`. Then the `renv` package will install all required packages with the correct version. After installing all packages, the `main_script.R` or individual scripts (`scripts/*`) can be used to run each analysis step.

## Folders organization

- `data/`: contains raw and cleaned data in `.rds` and `.xlsx`/`.csv` format
- `docs/`: contains scripts to create the supplementary materials
- `main_script.R`: is a script for easily managing all analysis steps
- `objects/`: contains R objects in `.rds` format created from the analysis scripts
- `R/`: contains all custom functions for the project. Functions are automatically loaded when the project is loaded. For reloading use `devtools::load_all()`
- `scripts/`: main scripts for pre-processing, statistical analysis and creating tables/figures
    - `1_analysis.R`: Scripts for calculating the effect size and computing the meta-analysis models
    - `2_tables_figures.R`: Script to create tables and figures for the paper
    - `3_supplementary.R`: Script to create supplementary materials objects
- `tables/`: contains tables created with the `2_tables_figures.R` in `.docx` format
- `tests/`: is for testing the effect size computation functions

## Coding style

The analysis project is organized as an R package where functions within the `R/` folder are automatically available into the global environment when the project is activated (`metaCoding.Rproj`). Is possible to manually load the functions using `devtools::load_all()`. The coding style is based on the [*tidyverse*](https://www.tidyverse.org/) using also [*metaprogramming*](https://adv-r.hadley.nz/meta-big-picture.html). For managing multiple meta-analysis models we used *nested tibbles* as data structure (see [here](https://r4ds.had.co.nz/many-models.html)). In particular the `objects/dat_meta.rds` contains all processing steps and models as different columns.

# Session Info

```{r, echo = FALSE}
sinfo <- sessioninfo::session_info()
sinfo$platform
```

# R Packages

```{r, results='asis', echo=FALSE}
pkgs <- sprintf("- `%s`", pkgs)
cat(pkgs, sep = "\n")
```

